### 2023-02-16 draft_manuscript.Rmd
Manuscript file - initial upload

---
title: "cellcount: an open-source and automated approach to cyanobacteria enumeration"
description: |
  blank text.
author:
  - name: Tyler E. Harman 
    affiliation: Consolidated Safety Services, Fairfax, VA 22031, USA
    affiliation_email: tyler.harman@noaa.gov
    orcid_id: 0000-0001-5007-8170
  - name: Mark W. Vandersea
    affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
  - name: David E. Berthold
    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
  - name: H. Dail Laughinghouse
    affiliation: Department of Biological Sciences, Florida International University, Miami, FL 33199, USA
    orcid_id: 0000-0003-1018-6948
  - name: Brandon Boyd
    affiliation: U.S. Army Corps of Engineers, Engineer Research and Development Center, Vicksburg, MS 39180, USA
  - name: Kaytee Pokrzywinski
    affiliation: National Oceanic and Atmospheric Administration, National Centers for Coastal Ocean Science, Beaufort, NC 28516, USA
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill for R Markdown at <https://rstudio.github.io/distill>.