### 2023-02-24 cellcount_GUI_intro.Rmd
Add files via upload

---
title: "cellcount - GUI format"
author: "T.E. Harman"
date: "2023-01-19"
abstract: >
  *cellcount*, a package available within the R environment, is an enumeration tool used to detect concentrations of cells within fluroescent microscopy images (see Harman et al. ____). This graphical user interface (GUI) takes from functions outlined in the *cellcount* package and applies them to a *shiny* interface for ease of use for end users, regardless of technological knowledge in R. This vignette demonstrates how to operate this package and provides examples for end-users to follow along.
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```